---
title: "Assignment4"
author: "Jason Chen"
date: "2024-11-16"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

# Completed {/10}

# 1

## a
```{r}
  set.seed(1)
  bird <- read.csv("NZBIRDS.csv")
  species <- bird[sample(1:nrow(bird), 35), ]
```

## b
```{r}
  mean <- mean(species$Body.Mass)
  mean
  sd <- sd(species$Body.Mass)
  sd
  t.test(bird$Body.Mass, mu = mean)
```

## c
The confidence interval shows that 95% of body mass would be in the interval (3208.217, 13397.434).

## d
The true mean, $\mu = 9113$, is included in the interval. The interval is likely to contain the true mean because it was created with a high probability (95%).

## e
```{r}
  mean <- mean(species$Egg.Length)
  mean
  sd <- sd(species$Egg.Length)
  sd
  t.test(bird$Egg.Length, mu = mean)
```

The confidence interval shows that 95% of egg length will be in the interval (58.02, 68.29).

The true mean, $\mu = 61.06$, is included in the interval. The interval is likely to contain the true mean because it was created with a high probability (95%).

## f
```{r}
  21/38 
  7/78
  qnorm(0.975)
  qnorm(0.975)*sqrt(0.0075)
```

$$
  (p_1-p_2) \pm Z_{\alpha-0.05} \sqrt{\frac{p_1(1-p_1)}{n_1}+\frac{p_2(1-p_2)}{n_2}} \newline
  = (0.5526-0.0897) \pm 1.96 \sqrt{\frac{0.5526(0.4474)}{38}+\frac{0.0897(0.9103)}{78}} \newline
  = 0.4629 \pm 1.96 \sqrt{0.0075} \newline
  = 0.4629 \pm 0.1697 \newline
  = [0.2932,0.6326]
$$

## g
The confidence interval supports this theory because the interval is positive meaning the ratio of extinct flightless birds is greater than the ratio of nonextinct flightless birds.

# 2

## a
```{r}
  qnorm(0.95)
```

$$
  (1312 - 1352) \pm 1.645 \sqrt{\frac{422^2}{100}+\frac{271^2}{47}} \newline
  = -40 \pm 1.645 \sqrt{1780.84+1562.57} \newline
  = -40 \pm 95.118 \newline
  = [-135.118, 55.118]
$$

## b
```{r}
  qf(0.95, 99, 46)
  qf(0.95, 46, 99)
```

$$
  \frac{422^2}{271^2}*\frac{1}{1.548} \leq \frac{\sigma_1^2}{\sigma_2^2} \leq
  \frac{422^2}{271^2}*{1.492}
  \newline = 1.566 \leq \frac{\sigma_1^2}{\sigma_2^2} \leq 3.618
$$

# 3

## a
$$
  Z = \frac{Y-\mu}{\sigma} = \frac{Y-0}{\sigma} = \frac{Y}{\sigma} \newline
$$
$Z^2 =  \frac{Y^2}{\sigma^2} \newline$ will have a chi-square distribution with 1 degree of freedom.

## b
$$
  X^2_{1-\alpha/2} \leq \frac{Y^2}{\sigma^2} \leq X^2_{\alpha/2} \newline
  = \frac{1}{X^2_{1-\alpha/2}} \geq \frac{\sigma^2}{Y^2} \geq \frac{1}{X^2_{\alpha/2}} \newline
  = \frac{1}{X^2_{\alpha/2}} \leq \frac{\sigma^2}{Y^2} \leq \frac{1}{X^2_{1-\alpha/2}} \newline
  = \frac{Y^2}{X^2_{\alpha/2}} \leq \sigma^2 \leq \frac{Y^2}{X^2_{1-\alpha/2}}
$$

# 4

## a
$$
  H_0: \mu=2 \newline
  H_a: \mu \neq 2
$$

## b
Test statistic = -1.02. p-value = 0.322.

## c
The rejection region for this test is $t < -1.96$ and $t > 1.96$ because $t_{0.05/2}=1.96$.

## d
We fail to reject the null hypothesis.

## e
The confidence interval and test statistic lead to the same conclusion about $\mu$ because both are methods to make inferences about a hypothesized value.

# 5

## a
```{r}
  lake <- read.csv("WISCLAKES.csv")
  mean(lake$DOC)
  sd(lake$DOC)
  qt(1-0.10/2, 25-1)
```

$$
  \frac{14.516-15}{\frac{12.964}{\sqrt25}} \newline
  = \frac{-0.484}{2.592} \newline
  = -0.187
$$

## b
$$
  \mu \pm t(\frac{s}{\sqrt n}) = 15 \pm 1.71 (\frac{12.96}{\sqrt 25}) \newline
  = 15 \pm 1.71 (2.59) \newline
  = [10.57, 19.43] \newline
  
  \frac{10.57-14}{2.59}= -1.32 \newline
  \frac{19.43-14}{2.59}= 2.1
$$

```{r}
  pt(-1.32,24)
  1-pt(2.1,24)
```
$$
  P(t<-1.32)+(t>2.1) = 0.099 + 0.023 = 0.122
$$

# 6

## a
```{r}
  orch <- read.csv("ORCHARD.csv")
  library(dplyr)
  dfog <- with(orch, orch[CONDITION == "FOG",])
  dfclcloud <- with(orch, orch[CONDITION == "CLEAR" | CONDITION == "CLOUD",])
  dfog
  dfclcloud
  F_RATIO <- subset(orch,CONDITION == "FOG", RATIO)
  CC_RATIO <- subset(orch,CONDITION == "CLOUD" | CONDITION == "CLEAR", RATIO)
  orch <- mutate(orch, Cond = recode(CONDITION, CLOUD = "CLCL", CLEAR = "CLCL"))
  orch
  
  sd(na.omit(F_RATIO)$RATIO)
  sd(CC_RATIO$RATIO)
```

$$
  s^2_r = \frac{(7)0.1186^2 + (3)0.1865^2}{8+4-2} = 0.02028
$$

```{r}
  mean(na.omit(F_RATIO)$RATIO)
  mean(CC_RATIO$RATIO)
```

$$
  \frac {(0.2738 - 0.4523) - 0}{\sqrt{0.02028(1/8+1/4)}} \newline
  = -0.1785 / 0.0883
  = -2.021
$$

# 7

## a
```{r}
  gas <- read.csv("GASTURBINE.csv")
  head(gas)
  trad <- filter(gas, ENGINE == "Traditional")
  aero <- filter(gas, ENGINE == "Aeroderiv")
  var.test(trad$HEATRATE, aero$HEATRATE)
  
  qf(c(0.05/2, 1-0.05/2), 6, 38)
  2652^2/1279^2
```

There is enough of a difference between traditional and aeroderivative engines' mean heat rates. 

## b
```{r}
  adv <- filter(gas, ENGINE == "Advanced")
  var.test(adv$HEATRATE, aero$HEATRATE)
```

There is enough of a difference between advanced and aeroderivative engines' mean heat rates.

# 8

## a
$$
  H_0: \frac{\sigma^2_1}{\sigma^2_2}=1 \newline
  H_a: \frac{\sigma^2_1}{\sigma^2_2} \neq 1
$$

## b
```{r}
  gobi <- read.csv("GOBIANTS.csv")
  head(gobi)
  dry <- filter(gobi, Region == "Dry Steppe")
  desert <- filter(gobi, Region == "Gobi Desert")
  length(dry)
  length(desert)
  sd1 = sd(dry$AntSpecies)
  sd1
  sd2 = sd(desert$AntSpecies)
  sd2
```

The test statistic is $\frac{21.307^2}{18.214^2} = 1.368$

## c
```{r}
  qf(c(0.05/2,1-0.05/2),6,6)
```

## d
```{r}
  var.test(dry$AntSpecies, desert$AntSpecies)
```

p-value of the test is 0.7264

## e
The variance of the populations at the two sites are not the same.

## f
Conditions that are required to make the test results valid are samples are independent, large samples to prevent deviations, and the true variance of the populations are equal.

# 9

$$
  H_0: \mu_1 -\mu_2 = 0 \newline
  H_a: \mu_1 -\mu_2 \neq 0
$$
```{r}
  thru <- read.csv("THRUPUT.csv")
  human <- thru$HUMAN
  auto <- thru$AUTO
  (mean(human)-mean(auto))/((sd(human-auto))/sqrt(length(human)))
  qt(c(0.05/2,1-0.05/2), 7)
```

The test statistic for the data lies outside the rejection region. So we reject the NULL Hypothesis that the mean time between humans and the automated method is equal.

# 10
```{r}
set.seed(35); sam<-round(rnorm(30,mean=20,sd=3),3)

myboot<-function(iter=10000,x,fun="mean",alpha=0.05,...){  #Notice where the ... is repeated in the code
n=length(x)   #sample size

y=sample(x,n*iter,replace=TRUE)
rs.mat=matrix(y,nr=n,nc=iter,byrow=TRUE)
xstat=apply(rs.mat,2,fun) # xstat is a vector and will have iter values in it 
ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
# A histogram follows
# The object para will contain the parameters used to make the histogram
para=hist(xstat,freq=FALSE,las=1,
main=paste("Histogram of Bootstrap sample statistics","\n","alpha=",alpha," iter=",iter,sep=""),col = rainbow(n),...)

# t-statistic for mean difference from 0
t_stat <- qt(1-alpha/2, df=n-1)

cit <- c(mean(x) - t_stat * (sd(x) / sqrt(n)), mean(x) + t_stat * (sd(x) / sqrt(n)))

#mat will be a matrix that contains the data, this is done so that I can use apply()
mat=matrix(x,nr=length(x),nc=1,byrow=TRUE)

# Theoretical confidence interval (blue line)
height1 <- max(para$density) * 0.2
text(cit[1], height1, paste("(", round(cit[1], 2), sep = ""), col = "blue", cex = 2)
text(cit[2], height1, paste(round(cit[2], 2), ")", sep = ""), col = "blue", cex = 2)

#pte is the point estimate
#This uses whatever fun is
pte=apply(mat,2,fun)
height2 <- max(para$density) * 0.1
abline(v=pte,lwd=3,col="Black")# Vertical line
segments(ci[1],0,ci[2],0,lwd=4)      #Make the segment for the ci
text(ci[1],height2,paste("(",round(ci[1],2),sep=""),col="Red",cex=2)
text(ci[2],height2,paste(round(ci[2],2),")",sep=""),col="Red",cex=2)

# plot the point estimate 1/2 way up the density
text(pte,max(para$density)/2,round(pte,2),cex=3)

return(list(fun=fun,x=x,t=t_stat,ci = ci,cit = cit))# Some output to use if necessary
}

myboot(x=sam)
```
